---
title: "Regression coefficients for interaction analyses"
output:
  html_document:
    df_print: paged
---

This notebook contains the regression coefficients for the interaction analyses (across all ancestries and all allele coding strategies).

The results of this notebook are also available in Supplementary Table 13 to 20

# Setup

```{r setup, include = FALSE}

library(dplyr)
library(purrr)
library(tidyr)
library(gt)
library(stringr)
library(gtExtras)

# Function for subgroups table

display_subgroups_table <- function(){
  c("rs1535", "rs174575", "rs174583") %>%
    map(
      \(x){
        summary_stats_plots %>%
          ungroup() %>%
          mutate(variant = str_extract(variant, "rs[0-9]+"),
                 lower_ci = estimate-(1.96*std.error),
                 upper_ci = estimate+(1.96*std.error)
          ) %>%
          filter(variant == x) %>%
          select(-c(variant, statistic, std.error)) %>%
          pivot_wider(
            names_from = stratum, 
            values_from = c(estimate, p.value, lower_ci, upper_ci)
          ) %>%
          gt(rowname_col = "covariates", groupname_col = "outcome") %>%
          sub_missing(
            columns = ends_with(c("Yes", "No", "Missing"))
          ) %>%
          fmt_number(
            columns = starts_with(c("estimate", "std.error", "lower_ci", "upper_ci")),
            decimals = 2
          ) %>%
          cols_merge(
            columns = c("estimate_Total",	"lower_ci_Total", "upper_ci_Total"),
            pattern = "<<{1}>><< [{2}, {3}]>>"
          ) %>%
          cols_merge(
            columns = c("estimate_Yes",	"lower_ci_Yes", "upper_ci_Yes"),
            pattern = "<<{1}>><< [{2}, {3}]>>"
          ) %>%
          cols_merge(
            columns = c("estimate_No",	"lower_ci_No", "upper_ci_No"),
            pattern = "<<{1}>><< [{2}, {3}]>>"
          ) %>%
          cols_merge(
            columns = c("estimate_Missing",	"lower_ci_Missing", "upper_ci_Missing"),
            pattern = "<<{1}>><< [{2}, {3}]>>"
          ) %>%
          tab_spanner(
            label = "Total",
            columns = (c("estimate_Total", "p.value_Total"))
          ) %>%
          tab_spanner(
            label = "Breastfed",
            columns = (c("estimate_Yes", "p.value_Yes"))
          ) %>%
          tab_spanner(
            label = "Bottlefed",
            columns = (c("estimate_No", "p.value_No"))
          ) %>%
          tab_spanner(
            label = "Missing",
            columns = (c("estimate_Missing", "p.value_Missing"))
          ) %>%
          cols_align(
            align = "center",
            columns = starts_with(c("estimate", "p.value"))
          ) %>%
          cols_label(
            estimate_Total = md("*β* [CI]"),
            p.value_Total = "p-value",
            estimate_Yes = md("*β* [CI]"),
            p.value_Yes = "p-value",
            estimate_No = md("*β* [CI]"),
            p.value_No = "p-value",
            estimate_Missing = md("*β* [CI]"),
            p.value_Missing = "p-value"
          ) %>%
          cols_align(align = "left", columns = 1)
      }
    ) %>%
    purrr::iwalk(\(table, SNP) table %>% tab_caption(SNP))
}

display_interaction_table <- function(){
  model_summary_interactions %>%
    filter(str_detect(term, "rs[0-9]+_[ACGT]+:breastfed_as_a_baby")) %>%
    mutate(variant = str_extract(term, "rs[0-9]+"),
           lower_ci = estimate-(1.96*std.error),
           upper_ci = estimate+(1.96*std.error)
    ) %>%
    select(-c(term, statistic, std.error)) %>%
    pivot_wider(
      names_from = variant, 
      values_from = c(estimate, p.value, lower_ci, upper_ci)
    ) %>%
    gt(rowname_col = "covariates", groupname_col = "outcome") %>%
    fmt_number(
      columns = starts_with(c("estimate", "std.error", "lower_ci", "upper_ci")),
      decimals = 2
    ) %>%
    cols_merge(
      columns = c("estimate_rs1535",	"lower_ci_rs1535", "upper_ci_rs1535"),
      pattern = "<<{1}>><< [{2}, {3}]>>"
    ) %>%
    cols_merge(
      columns = c("estimate_rs174575",	"lower_ci_rs174575", "upper_ci_rs174575"),
      pattern = "<<{1}>><< [{2}, {3}]>>"
    ) %>%
    cols_merge(
      columns = c("estimate_rs174583",	"lower_ci_rs174583", "upper_ci_rs174583"),
      pattern = "<<{1}>><< [{2}, {3}]>>"
    ) %>%
    tab_spanner(
      label = "rs1535",
      columns = (c("estimate_rs1535", "p.value_rs1535"))
    ) %>%
    tab_spanner(
      label = "rs174575",
      columns = (c("estimate_rs174575", "p.value_rs174575"))
    ) %>%
    tab_spanner(
      label = "rs174583",
      columns = (c("estimate_rs174583", "p.value_rs174583"))
    ) %>%
    cols_align(
      align = "center",
      columns = starts_with(c("estimate", "p.value"))
    ) %>%
    cols_label(
      estimate_rs1535 = md("*β* [CI]"),
      p.value_rs1535 = "p-value",
      estimate_rs174575 = md("*β* [CI]"),
      p.value_rs174575 = "p-value",
      estimate_rs174583 = md("*β* [CI]"),
      p.value_rs174583 = "p-value"
    ) %>%
    cols_align(align = "left", columns = 1)
}
```

# Additive coding

## White British

### Subgroups

```{r white-british-all, include = FALSE, results="asis"}

base::load(snakemake@input[[1]])

display_subgroups_table()
```

### Interaction terms

```{r interaction-additive-white-british, include = FALSE}
display_interaction_table()
```

## Any ancestry

### Subgroups

```{r all-ancestries-additive, include = FALSE, results = "asis"}

base::load(snakemake@input[[2]])

display_subgroups_table()
```

### Interaction terms

```{r interaction-additive-all, include = FALSE}
display_interaction_table()
```

# Recessive coding

## White British

### Subgroups

```{r white-british-recessive, include = FALSE, results = "asis"}
base::load(snakemake@input[[3]])

display_subgroups_table()
```

### Interaction terms

```{r interaction-recessive-white-british, include = FALSE}

display_interaction_table()
```

## Any ancestry

### Subgroups

```{r all-ancestries-recessive, include = FALSE, results = "asis"}
base::load(snakemake@input[[4]])

display_subgroups_table()
```

### Interaction terms

```{r interaction-recessive-any-ancestry, include = FALSE}
display_interaction_table()
```
