---
title: "Phenotypic Analysis"
author: "Giulio Centorame"
format: html
editor: visual
---

This notebook contains the phenotypic association analysis between breastfeeding and our outcomes of interest

```{r setup, include = FALSE}

library(dplyr)
library(purrr)
library(tidyr)
library(gtsummary)
library(modelsummary)

# Load data in
base::load(snakemake@input[[1]])

descriptive_data <-
  # Round dosage to allelic count
  descriptive_data %>%
  mutate(
    # Switch to implicit NAs 
    breastfed_as_a_baby = na_if(breastfed_as_a_baby, "Missing"),
    breastfed_as_a_baby = recode(
      breastfed_as_a_baby,
      Yes = 1,
      No = 0
      ),
    sex = recode(sex, Female = 0, Male = 1)
    ) %>%
  rename(yob = year_of_birth_f34_0_0,
         tdi = townsend_deprivation_index_at_recruitment_f189_0_0)

rename_list <- function(.data, ...) {
    mapping <- sapply(
        rlang::enquos(...),
        rlang::as_name
    )
    new_names = setNames(nm=names(.data))
    # `new_name = old_name` for consistency with `dplyr::rename`
    new_names[mapping] = names(mapping)
    # for `old_name = new_name` use: `new_names[names(mapping)] = mapping`
    setNames(.data, new_names)
}

# Create character vector for formula
formula_bf <-
  function(outcome, delivery = FALSE){
    
    basic_formula <- paste(outcome, "~ breastfed_as_a_baby")
    
    if(delivery){
      
      return(paste(basic_formula, "+ delivery"))  
      
    } else {
      
      return(basic_formula)
    }
  }

formula_bf_age_sex <-
  function(outcome, delivery = FALSE){
    paste(formula_bf(outcome = outcome, delivery = delivery), "+ age + sex")
  }

formula_bf_age_sex_tdi <-
  function(outcome, delivery = FALSE){
    paste0(formula_bf_age_sex(outcome = outcome, delivery = delivery), "+ tdi")
  }

formula_bf_age_sex_tdi_pcs <-
  function(outcome, delivery = FALSE){
    paste(formula_bf_age_sex_tdi(outcome = outcome, delivery = delivery), 
          paste("+", paste0("PC", 1:40), collapse = " + "))
  }

# Add descriptive stats to data

list_continuous_single_delivery_first_instance_breastfeeding <-
  list_continuous_single_delivery_first_instance_breastfeeding %>%
  map(\(x) inner_join(x, descriptive_data, by = "eid"))

list_continuous_multiple_delivery_first_instance_breastfeeding <-
  within(list_continuous_multiple_delivery_first_instance_breastfeeding, rm(pairs_matching_incorrect_matches))

list_continuous_multiple_delivery_first_instance_breastfeeding <-
  list_continuous_multiple_delivery_first_instance_breastfeeding %>%
  map(\(x) inner_join(x, descriptive_data, by = "eid"))

list_binary_first_instance_breastfeeding <-
  list_binary_first_instance_breastfeeding %>%
  map(\(x) inner_join(x, descriptive_data, by = "eid")) %>% 
  within(., rm(prospective_memory_binary))

rm(common_data, descriptive_data)
```

# Variable coding

-   Breastfeeding is coded as 0 = bottlefed, 1 = breastfed

-   Sex is coded as 0 = Female, 1 = Male

-   Delivery is coded as Assessment centre = 0, Online = 1

# Association with breastfeeding

## Continuous traits

```{r continuous-single-delivery, echo = FALSE}

cont_sd_unadj <-
 list_continuous_single_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf(trait)), data = data)
       )

cont_sd_age_sex <-
 list_continuous_single_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf_age_sex(trait)), data = data)
       )

cont_sd_age_sex_tdi <-
 list_continuous_single_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf_age_sex_tdi(trait)), data = data)
       )

cont_sd_age_sex_tdi_pcs <-
 list_continuous_single_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf_age_sex_tdi_pcs(trait)), data = data)
       )

# Remove unused object to save some memory

rm(list_continuous_single_delivery_first_instance_breastfeeding)

list(cont_sd_unadj, 
     cont_sd_age_sex, 
     cont_sd_age_sex_tdi, 
     cont_sd_age_sex_tdi_pcs) %>%
  map(
    \(x)
    rename_list(
      x,
     "Years of schooling" = years_of_schooling,
     "Number of offspring" = number_of_offspring,
     "Total cholesterol" = total_cholesterol,
     "LDL cholesterol (mmol/l)" = ldl_direct,
     "HDL cholesterol (mmol/l)" = hdl_direct,
     "Triglycerides (mmol/l)" = triglycerides,
     "Body mass index (kg/m²)" = bmi,
     "Systolic blood pressure (mmHg, controlling for medication use)" = sbp,
     "Diastolic blood pressure (mmHg, controlling for medication use)" = dbp,
     "Matrix pattern completion (correct answers)" = correct_answers_matrix_completion,
     "Reaction time (ms)" = reaction_time_ms,
     "Tower rearranging (correct answers)" = tower_rearranging_correct_answers #,
     #"Paired associate learning (correct word pairs)" =  paired_associate_learning_correct_word_pairs
     )
    ) %>%
  pmap(
  \(unadj, agesex, tdi, pcs) {
    list(
      "Unadjusted" = unadj,
      "Adjusted for age and sex" = agesex,
      "Adjusted for age, sex and TDI" = tdi,
      "Adjusted for age, sex, TDI, and PCs" = pcs
      )
}) %>%
  modelsummary(
    shape = "rbind",
        statistic = 'conf.int',
        coef_rename = c("age" = "Age",
                        "sex" = "Sex",
                        "breastfed_as_a_baby" = "Breastfed as a baby",
                        "tdi" = "Townsend deprivation index",
                        "deliveryOnline" = "Delivery mode"))

# Free some more memory
rm(cont_sd_unadj, 
   cont_sd_age_sex, 
   cont_sd_age_sex_tdi, 
   cont_sd_age_sex_tdi_pcs)
```

## Cognitive traits with multiple delivery mode

```{r continuous-multiple-delivery, echo = FALSE}

cont_md_unadj <-
 list_continuous_multiple_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf(trait, delivery = TRUE)), data = data)
       )

cont_md_age_sex <-
 list_continuous_multiple_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf_age_sex(trait, delivery = TRUE)), data = data)
       )

cont_md_age_sex_tdi <-
 list_continuous_multiple_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf_age_sex_tdi(trait, delivery = TRUE)), data = data)
       )

cont_md_age_sex_tdi_pcs <-
 list_continuous_multiple_delivery_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      lm(formula(formula_bf_age_sex_tdi_pcs(trait, delivery = TRUE)), data = data)
       )

# Free some memory

rm(list_continuous_multiple_delivery_first_instance_breastfeeding)

list(cont_md_unadj,
     cont_md_age_sex, 
     cont_md_age_sex_tdi,
     cont_md_age_sex_tdi_pcs) %>%
  map(
    \(x)
    rename_list(
      x,
       "Fluid intelligence" = fluid_intelligence,
  "Numeric memory (max digits)" = numeric_memory_max_digits,
  "Symbol digit substitution (correct answers)" = symbol_digit_substitution_correct_answers,
  "Trailmaking path 1 (ds)" = trailmaking_path_1,
  "Trailmaking path 2 (ds)" = trailmaking_path_2,
  "Trailmaking path 2-1 (ds)" = trailmaking_2_minus_1
     )
    ) %>%
  pmap(
  \(unadj, agesex, tdi, pcs) {
    list(
      "Unadjusted" = unadj,
      "Adjusted for age and sex" = agesex,
      "Adjusted for age, sex and TDI" = tdi,
      "Adjusted for age, sex, TDI, and PCs" = pcs
      )
}) %>%
  modelsummary(
    shape = "rbind",
        statistic = 'conf.int',
        coef_rename = c("age" = "Age",
                        "sex" = "Sex",
                        "breastfed_as_a_baby" = "Breastfed as a baby",
                        "tdi" = "Townsend deprivation index",
                        "deliveryOnline" = "Delivery mode"))

# Free even more memory

rm(cont_md_unadj,
   cont_md_age_sex, 
   cont_md_age_sex_tdi,
   cont_md_age_sex_tdi_pcs)
```

## Binary traits

```{r binary-traits, echo = FALSE}

bin_unadj <-
 list_binary_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      glm(formula(formula_bf(trait)), data = data, family=binomial(link="logit"))
       )

bin_age_sex <-
 list_binary_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      glm(formula(formula_bf_age_sex(trait)), data = data, family=binomial(link="logit"))
       )

bin_age_sex_tdi <-
 list_binary_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      glm(formula(formula_bf_age_sex_tdi(trait)), data = data, family=binomial(link="logit"))
       )

bin_age_sex_tdi_pcs <-
 list_binary_first_instance_breastfeeding %>%
  imap(
    \(data, trait)
      glm(formula(formula_bf_age_sex_tdi_pcs(trait)), data = data, family=binomial(link="logit"))
       )

# Free some memory

rm(list_binary_first_instance_breastfeeding)

list(bin_unadj,
     bin_age_sex,
     bin_age_sex_tdi,
     bin_age_sex_tdi_pcs) %>%
  map(
    \(x)
    rename_list(
      x,
      "Angina" = angina,
    "Heart attack" = heart_attack,
    "Stroke" = stroke,
    "Type-II Diabetes" = type2_diabetes,
    "Hayfever, rhinitis, or eczema" = hayfever_rhinitis_eczema
     )
    ) %>%
  pmap(
  \(unadj, agesex, tdi, pcs) {
    list(
      "Unadjusted" = unadj,
      "Adjusted for age and sex" = agesex,
      "Adjusted for age, sex and TDI" = tdi,
      "Adjusted for age, sex, TDI, and PCs" = pcs
      )
}) %>%
  modelsummary(
    shape = "rbind",
        statistic = 'conf.int',
        coef_rename = c("age" = "Age",
                        "sex" = "Sex",
                        "breastfed_as_a_baby" = "Breastfed as a baby",
                        "tdi" = "Townsend deprivation index",
                        "delivery" = "Delivery mode"))

# Free some more memory

rm(bin_unadj,
   bin_age_sex,
   bin_age_sex_tdi,
   bin_age_sex_tdi_pcs)
```
